---
title: "ARID1A DepMap data Analysis"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---



```{r}
library(tidyverse)
library(reshape2)
library(ggbiplot)
library(scales)
library(svglite)
library(factoextra)
library(ggsci)
library(Rtsne)
library(ggrepel)
library(ggpubr)


```
```{r}
theme_su<-  function(){ 
  font <- "Arial"   #assign font family up front
  
  theme_bw() %+replace%    #replace elements we want to change
    
    theme(
   
      
      #text elements
      plot.title = element_text(             #title
        family = "Arial",            #set font family
        size = 8,                #set font size
        face = 'bold',            #bold typeface
        hjust = 0,                #left align
        vjust = 3),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
        family = "Arial",            #font family
        size = 6),               #font size
      
      plot.caption = element_text(           #caption
        family = "Arial",            #font family
        size = 6,                 #font size
        hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
        family = "Arial",
        
        size = 6),               #font size
      
      axis.text = element_text(              #axis text
        family = "Arial",            #axis famuly
        size = 6),                #font size
      
      axis.text.x = element_text(            #margin for axis text
        margin=margin(5, b = 10)),
      
     legend.title=element_text(
       family="Arial",
       size=8,
       hjust = 0),
     legend.text=element_text(
       family = "Arial",
       size=6),
     
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank()
     )
  
     
    
}
```


### data acquisition and cleaning
Depmap data and cell lines data are downloaded from depmap database (December, 2021)

```{r}
arid1a_mutations<- read.csv(file="input/ARID1A mutations.csv")
print(paste("there are", nrow(arid1a_mutations), "mutations in ARID1A gene",length(unique(arid1a_mutations$Depmap.Id)), "lines reported in Depmap database "))

table(arid1a_mutations[,11:12])
```


```{r}
CCLE_mutations <- read.csv("input/CCLE_mutations.csv")

print(paste("there are", nrow(CCLE_mutations), "mutations in ",length(unique(CCLE_mutations$DepMap_ID)), "lines reported in Depmap database "))
```
#I removed silent mutations
```{r}
arid1a_mutations_damaging<- arid1a_mutations %>% filter(Variant.Annotation !="silent") %>% droplevels()
print(paste("there are", nrow(arid1a_mutations_damaging), "damaging mutations in ARID1A gene",length(unique(arid1a_mutations_damaging$Depmap.Id)), "lines reported in Depmap database "))
```

```{r}
CCLE_mutations_damaging<- CCLE_mutations %>% filter(isDeleterious =="True") %>% droplevels()
print(paste("there are", nrow(CCLE_mutations_damaging), "damaging mutations in",length(unique(CCLE_mutations_damaging$DepMap_ID)), "lines reported in Depmap database and in", length(unique(CCLE_mutations_damaging$Hugo_Symbol)), "Genes"))

```


```{r}
number_of_mutations_per_line<- as.data.frame(table(CCLE_mutations_damaging$DepMap_ID))
lines_with_arid1a_mutattion<- CCLE_mutations_damaging %>% filter(Hugo_Symbol=="ARID1A") %>% select(DepMap_ID) %>% unique()

lines_with_p53_mutation<- CCLE_mutations_damaging %>% filter(Hugo_Symbol=="TP53") %>% select(DepMap_ID) %>% unique()

a<- as.data.frame(table(CCLE_mutations_damaging[,c("Hugo_Symbol", "DepMap_ID")]))
a<- a %>% filter(Freq!=0)
b<- as.data.frame(table(a$Hugo_Symbol))
b %>% ggplot(aes(y=Freq)) +geom_boxplot()
b<-b %>% filter(Var1!="TTN") %>% droplevels()

b$label <- paste(b$Var1, b$Freq, sep= "-")

sup.fig.8A<-b %>% ggplot(aes(y=log(Freq,10),x=1))+
  geom_boxplot(fill="lightblue", outlier.size = 0.5)+
  geom_label_repel(data=(b %>% filter(Freq>150)), aes(label=label),min.segment.length = 0, size=2, nudge_x = T)+
  theme(legend.position="none")+ylab("Number of mutated cell lines (Log2)")+theme_su()+theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank())
sup.fig.8A

ggsave(sup.fig.8A,filename ="output/sup_fig_8A.svg", width = 2, height = 2, units = "in", dpi=1000 )


number_of_mutations_per_line<- number_of_mutations_per_line %>% mutate(status=case_when(
  Var1 %in% lines_with_arid1a_mutattion$DepMap_ID ~"ARID1A",
  TRUE ~"Other"
))


number_of_mutations_per_line<- number_of_mutations_per_line %>% mutate(status_2=case_when(
  Var1 %in% lines_with_arid1a_mutattion$DepMap_ID ~"ARID1A",
  Var1 %in% lines_with_p53_mutation$DepMap_ID~ "P53",
  TRUE ~"Other"
))

number_of_mutations_per_line<- number_of_mutations_per_line %>% mutate(status_3=case_when(
  Var1 %in% lines_with_arid1a_mutattion$DepMap_ID &Var1 %in% lines_with_p53_mutation$DepMap_ID ~"Both",
  Var1 %in% lines_with_arid1a_mutattion$DepMap_ID ~"ARID1A",
  Var1 %in% lines_with_p53_mutation$DepMap_ID~ "P53",
  TRUE ~"Other"
))

number_of_mutations_per_line<- do.call("data.frame", number_of_mutations_per_line)
table(number_of_mutations_per_line$status_3)

number_of_mutations_per_line %>% ggplot(aes(x=status_3, y=Freq))+geom_boxplot()+ylab("Number of mutation per line")


x<- CCLE_mutations_damaging %>% filter(DepMap_ID %in% lines_with_arid1a_mutattion$DepMap_ID) %>% droplevels()

median(number_of_mutations_per_line$Freq)

n<- number_of_mutations_per_line %>% dplyr::group_by(status) %>% dplyr::summarise(median=median(Freq))



```




arid1a is among the most mutated lines.
check how many different genes are mutated together with arid1a







find no damaging mutation lines
```{r}
arid1a_lines_with_no_damaging_mutation<- arid1a_mutations %>% filter(!(Cell.Line %in% arid1a_mutations_damaging$Cell.Line)) %>% droplevels()
cell_line_data<-as.data.frame(unique(arid1a_mutations$Depmap.Id))
names(cell_line_data)<- "DepMap_ID"
cell_line_data$mutation_type<- "Damaging"
cell_line_data<- cell_line_data %>% mutate(mutation_type=case_when(
  DepMap_ID %in% arid1a_lines_with_no_damaging_mutation$Depmap.Id ~"Benign",
    TRUE ~ as.character(mutation_type)))
```


### import screen data

```{r}
arid1a_screens<- read.csv(file="input/arid1a_screens.csv")

arid1a_screens<- left_join(arid1a_screens, cell_line_data, by="DepMap_ID")
arid1a_screens<- arid1a_screens %>% select(1,17388, 2:17387)
#also load data for all screens
CRISPR_gene_effect <- read.csv("input/CRISPR_gene_effect.csv")

##import DDD bagel file
DDD_Bagel <- read.csv("input/DDD_Bagel.txt")
```



#I used bagel values because bagel is also normalizes screens according to scores of essential and non essential genes.


```{r}
names(DDD_Bagel)
DDD_bagel_screen<- DDD_Bagel %>% select(Gene, WT_1, WT_2, WT_av, Mutant)
DDD_bagel_screen<-DDD_bagel_screen %>% mutate(WT_1= -1*WT_1, WT_2=-1*WT_2, WT_av= -1*WT_av, Mutant= -1* Mutant )
rownames(DDD_bagel_screen)<- DDD_bagel_screen$Gene
DDD_long<- as.data.frame(t(DDD_bagel_screen[,2:5]))

```
##combine screen data with ddd data

```{r}
c<- colnames(CRISPR_gene_effect)
c<- gsub(pattern = "\\..*", replacement = "",c)
names(CRISPR_gene_effect)<- c
DDD_long$DepMap_ID<- rownames(DDD_long)
t<- intersect(names(DDD_long),c)

##only 16626 genes are shared between data sets

m<- rbind(CRISPR_gene_effect[,t], DDD_long[,t])
names(m)[1]
rownames(m)<- m$DepMap_ID
m<- m %>% mutate(screen_type="Other") %>% mutate(case_when(
    DepMap_ID %in% DDD_long$DepMap_ID ~ "in_house",
  DepMap_ID %in% lines_with_arid1a_mutattion ~"Depmap_ARID1A",
  TRUE ~ as.character(screen_type)))
all_screen_data<-m
```


```{r}
n<- as.data.frame(t(m[1:16628]))

```

##Correlation plot
```{r}
cormat<- round(cor(n, method = "pearson"),3)


melted_cormat <- melt(cormat)
head(melted_cormat)


a<- melted_cormat %>% filter((Var1 %in% c("WT_av", "Mutant"))&!(Var2 %in% c("WT_av", "Mutant", "WT_1","WT_2"))) %>% droplevels()
b<-as.data.frame(acast(a, Var2~Var1))
b %>% ggplot(aes(x=WT_av, y=Mutant))+geom_point()+geom_abline()
library(ggpubr)


a<- a %>% mutate(screen_type="Other") %>% mutate(screen_type=case_when(
    Var2 %in% lines_with_arid1a_mutattion$DepMap_ID ~"Depmap_ARID1A",
  TRUE ~ as.character(screen_type)))

correlations<- a
sup.fig.8C<-correlations %>% ggplot(aes(y=value, x=Var1))+geom_boxplot(outlier.shape =NA)+stat_compare_means(label="..p.signif..", label.y = 0.67)+theme_su()+ggtitle("Correlation between in house screens to all Depmap screens")+ xlab("in house screens")+ylab("Pearson Correlation")+ylim(0.5,0.7)
sup.fig.8C

ggsave(sup.fig.8C, filename = "output/sup_fig_8C.svg", width = 2.5, height = 2.5, units = "in", dpi=1000)

sup.fig.8D<-correlations %>% ggplot(aes(y=value, x=Var1, fill=screen_type))+geom_boxplot(outlier.shape = NA)+stat_compare_means(label="..p.signif..", size=2, label.y = 0.67)+theme_su()+ scale_fill_npg(name="Depmap Screen Type")+ ggtitle("Comparison of in-house screens to Depmap screens")+xlab("In-house screens")+ylab("Pearson correlations")+ylim(0.5,0.7)
sup.fig.8D

ggsave(sup.fig.8D, filename = "output/sup_fig_8D.svg", width = 4, height = 2.5, units = "in", dpi=1000)

```


###PCA analysis
```{r}
all_screen_data<- na.omit(all_screen_data)
depmap_screens_only<- CRISPR_gene_effect[,t] %>% na.omit()

head(depmap_screens_only)

depmap_screens_only<-depmap_screens_only %>% mutate(screen_type="Other") %>% mutate(screen_type=case_when(
  DepMap_ID %in% lines_with_arid1a_mutattion$DepMap_ID ~"Depmap_ARID1A",
  TRUE ~ as.character(screen_type)))


depmap.pca<- prcomp(depmap_screens_only[,1:16628], center = T, scale. = T)
library(ggfortify)
p1<-autoplot(depmap.pca, data = depmap_screens_only, colour = 'screen_type', size=0.1)
sup.fig.8B<- p1+theme_su()+scale_color_nejm()
sup.fig.8B
ggsave(sup.fig.8B, file="output/sup.fig.8B.svg", width=4, height = 2.5, units = "in", dpi=1000)

```
