---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

Bagel2 was used to analyse reads. BF value is a indicator of gene essentiality combined with significance. a higher BF value shows a higher depletion in screen.



```{r}
setwd("E:/Paper writing/CRIPSRi paper/Analyses/Dolcetto_screen_analyses/publication/publication_final")
library(tidyverse)
library(ggsci)
library(reshape2)
library(ineq)
library(ggpubr)
library(cowplot)
library(ggrepel)
library(svglite)
```


```{r}
theme_su<-  function(){ 
  font <- "Arial"   #assign font family up front
  
  theme_bw() %+replace%    #replace elements we want to change
    
    theme(
   
      
      #text elements
      plot.title = element_text(             #title
        family = "Arial",            #set font family
        size = 8,                #set font size
        face = 'bold',            #bold typeface
        hjust = 0,                #left align
        vjust = 3),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
        family = "Arial",            #font family
        size = 6),               #font size
      
      plot.caption = element_text(           #caption
        family = "Arial",            #font family
        size = 6,                 #font size
        hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
        family = "Arial",
        
        size = 6),               #font size
      
      axis.text = element_text(              #axis text
        family = "Arial",            #axis famuly
        size = 6),                #font size
      
      axis.text.x = element_text(            #margin for axis text
        margin=margin(5, b = 10)),
      
     legend.title=element_text(
       family="Arial",
       size=8,
       hjust = 0),
     legend.text=element_text(
       family = "Arial",
       size=6),
     
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank()
     )
  
     
    
}
```

### import gene level Bagel results for all screens

```{r}
all_Bagel <- read.csv("input/all_Bagel_curated.txt",sep=",")

```
##comparing monoclonal and polyclonal lines
```{r}

mono_vs_poly_bagel<- all_Bagel %>% select(Gene, func, IPSC_polyclonal= D1P_D22, IPSC_monoclonal=D1M_D18)

all_dolcetto_LFC_Gene<- read.csv( file="output/all_dolcetto_curated_LFC.txt")

LFC_density<-all_dolcetto_LFC_Gene %>% select(func, D1P_D22, D1M_D18)%>% filter(func %in% c("CEG", "NEG")) %>% melt()  %>% ggplot(aes(x=value, fill=variable), color="black")+geom_density(alpha=0.7)+
theme_su()+scale_fill_nejm(name="hiPSC + dCas9 Line", 
                                                                          labels=c("IPSC_Poly_D18"="Polyclonal" , "IPSC_D18"="Monoclonal"))+
 xlab("Gene Mean LFC")+ ggtitle("Gene Mean Log2 Fold Change")+ scale_linetype(name="Screen")
LFC_density


BF_density<-
mono_vs_poly_bagel %>% filter(func %in% c("CEG", "NEG")) %>% melt(id=c("Gene", "func")) %>% 
ggplot(aes(x=value,  fill=variable))+geom_density(color="black", alpha=0.7)+ theme_su()+scale_fill_nejm(name="hiPSC + dCas9 Line", 
                                                                          labels=c("IPSC_Poly_D18"="Polyclonal" , "IPSC_D18"="Monoclonal"))+ ggtitle("Bagel Analysis")+xlab("Bayes Factor (BF)")+theme(axis.title.y = element_blank())
BF_density

fig_3_f<-ggarrange(LFC_density, BF_density,nrow=1, common.legend = T, legend = "bottom" )

fig_3_f<- annotate_figure(fig_3_f, top = text_grob("Polyclonal vs Monoclonal hiPSC Screens Comparison", 
               face = "bold", size = 10))
fig_3_f



ggsave(fig_3_f, filename= "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_3/fig_3_f.svg", height = 3, width = 4,units="in", dpi=1000)

```
polyclonal and monoclonal lines shows different depletion pattern in gene mean log2 fold change but Bagel analysis normalizes the screens with different dynamics

next we decided to see if both screens capture same genes as essential genes. the intersection point to differentiate essential and non essential genes is around 6, which is als, so we compared genes with BF values above 0 between two screens:

```{r}
a<- mono_vs_poly_bagel %>% filter(IPSC_polyclonal>6) %>% select(Gene)
b<- mono_vs_poly_bagel %>% filter(IPSC_monoclonal>6) %>% select(Gene)

print(paste("# of genes common between two screens",length(intersect(a$Gene, b$Gene))))
print(paste("ratio of common genes over all genes is",round((length(intersect(a$Gene, b$Gene))/length(union(a$Gene, b$Gene))),2)))
```

### nocodazole test comparison:

```{r}
Noc_test<- all_Bagel %>% select(Gene, func, D2_Noc=D2M1R02_Noc, D2_Cont=D2M1R02_Cont)
names(Noc_test)
```

```{r}
Noc_test %>% ggplot(aes(x=D2_Noc, y=D2_Cont, color=func))+geom_point() 

Noc_test$difference<- Noc_test$D2_Cont - Noc_test$D2_Noc
Noc_test$Z_score<-scale(Noc_test$difference)
Noc_test %>% ggplot(aes(x=difference))+geom_density()
Noc_test_melted<- melt(Noc_test)

supp_fig_3_i<-Noc_test_melted %>% filter(variable %in% c("D2_Noc", "D2_Cont")) %>% ggplot(aes(y=value, x=variable, fill=variable))+ geom_boxplot(outlier.shape = NA)+stat_compare_means(label.y =30)+scale_fill_aaas(name="Treatment")+ ylim(-20, 35)+theme_su()+ theme(axis.title.x =element_blank())+ ylab("Bayes Factor")
supp_fig_3_i

ggsave(supp_fig_3_i, filename= "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_5/supp_fig_5_i.svg", height = 3, width = 4,units="in", dpi=1000)



Noc_test %>% ggplot(aes(x=Z_score))+geom_density()

## genes with treatment and control difference outside of 99% CI (Z score > 2.576 or Z score -2.576) are labeled as enriched and depleted respectively.

Noc_test<- Noc_test %>%  mutate(group="Normal") %>% mutate(group=case_when(
  Z_score > 2.576 ~ "Enriched",
  Z_score < -2.576 ~ "Depleted",
  TRUE ~as.character(group)
))
Noc_test<- do.call(data.frame, Noc_test)
summary(Noc_test$group)
```

```{r}
Noc_test$rank<- rank(Noc_test$difference)

fig_5_a<-Noc_test%>% ggplot(aes(x=rank, y=difference, color=group))+geom_point(size=0.2)+ geom_label_repel(data=(Noc_test %>% filter(rank >18869)),aes(x=rank, y=difference,label=Gene), size=2)+theme_su()+scale_color_nejm(name="Gene Silencing Effect\non Treatment")+ylab("BF Difference\n (Treatment vs Control)")+ xlab("Rank of Difference")+ ggtitle("Nocodazole Treatment Effect")
fig_5_a

ggsave(fig_5_a, filename= "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_5/fig_5_a_new.svg", height = 2, width = 3.5, units = "in", dpi=1000)


```


### similar approach is used to identify genes differentially essential in ARID1A mutant background

```{r}
DDD_bagel<- all_Bagel %>% select(Gene, func, WT_1=D2M1_All_D21, WT_2=D2M2_All_D18, Mutant=ARID1A_HT_All_D18)
DDD_bagel$WT_av<- (DDD_bagel$WT_1 + DDD_bagel$WT_2)/2
DDD_bagel$difference<- DDD_bagel$WT_av - DDD_bagel$Mutant
DDD_bagel %>% ggplot(aes(x=difference))+geom_density()+geom_vline(xintercept = mean(DDD_bagel$difference))+geom_vline(xintercept = (mean(DDD_bagel$difference)+2*sd(DDD_bagel$difference)))+geom_vline(xintercept = (mean(DDD_bagel$difference)-2*sd(DDD_bagel$difference)))

```



```{r}
DDD_melted_BF<- DDD_bagel %>% filter(func %in% c("CEG", "NEG")) %>% select(Gene, func, WT_av, Mutant) %>% melt
supp_fig_5_a<-DDD_melted_BF %>% ggplot(aes(x=value, color= func, linetype=variable))+geom_density()+ theme_su()+scale_color_nejm()
supp_fig_5_a
ggsave(supp_fig_5_a, filename ="E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_5/supp_fig_5_a_new.svg", width =3, height=3, units = "in", dpi=1000)

```




```{r}
## genes with treatment and control difference outside of 2sd  are labeled as enriched and depleted respectively.

DDD_bagel<- DDD_bagel %>%  mutate(group="Normal") %>% mutate(group=case_when(
  difference > mean(difference) + 2*sd(difference) ~ "Enriched",
  difference < mean(difference) - 2*sd(difference) ~ "Depleted",
  TRUE ~as.character(group)))
table(DDD_bagel$group)





scatter1<-
ggplot()+
  geom_point(data=(DDD_bagel %>% filter(func=="Other") %>% droplevels()), aes(x=WT_av, y= Mutant, color="Other"),size=0.5, alpha=0.6)+
  geom_point(data=(DDD_bagel %>% filter(func=="CEG") %>% droplevels()), aes(x=WT_av, y= Mutant, color="CEG"),size=0.5, alpha=0.8)+
  geom_point(data=(DDD_bagel %>% filter(func=="NEG") %>% droplevels()), aes(x=WT_av, y= Mutant, color="NEG"),size=0.5, alpha=0.8)+
ggtitle(expression(paste("Comparison of Screens in ARID1A" ^"+/-","and ARID1A"^"+/+", "Cells (BF Values)")))+ylab(expression(paste("ARID1A" ^"+/-","Cells")))+xlab(expression(paste("ARID1A" ^"+/+","Cells")))+theme_su()+ scale_color_nejm(name="Gene Function",labels=c("CEG"="Core Essential Genes" , "NEG"="Non-Essential Genes",  "Non-targeting_control"="Control", "Other"="Other Genes"))+theme(legend.position = "bottom")+ geom_label_repel(data=(DDD_bagel %>% filter(Gene %in% c("PSMB2", "NF2","AP1G1", "SEL1L", "EXOSC7", "TAF6L", "EZH2"))),aes(x=WT_av, y= Mutant, label=Gene, color=func), box.padding =2, size=2, max.overlaps = 10)

scatter1

write_csv(DDD_bagel, file = "E:/Paper writing/CRIPSRi paper/Analyses/Dolcetto_screen_analyses/publication/publication_final/output/DDD_bagel_with_difference.txt")



## then add correlation data for all data together (correlation without p value)  
scatter1<- scatter1+ stat_cor(data=DDD_bagel, aes(x=WT_av, y= Mutant, label = ..r.label..), inherit.aes = F, size=3)
scatter1
##then produce density plots separately and add them to edges of scatter plot
xdens <- axis_canvas(scatter1, axis = "x")+
  geom_density(data = DDD_bagel, aes(x = WT_av, fill = func),
              alpha = 0.7, size = 0.2)+
 scale_fill_nejm()
ydens <- axis_canvas(scatter1, axis = "y", coord_flip = T)+
  geom_density(data = DDD_bagel, aes(x = Mutant, fill = func),
              alpha = 0.7, size = 0.2)+
 scale_fill_nejm()+coord_flip()


p1 <- insert_xaxis_grob(scatter1, xdens, grid::unit(.2, "null"), position = "top")
p2<- insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = "right")
p3<- ggdraw(p2)
p3
ggsave(p3, filename = "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_5/fig_5_b_new_new.svg", width = 2.5, height = 3, units = "in", dpi=1000)



```

### I will conduct similar analyses with gene mean LFC count


```{r}
##derive final timepoint all reads summed LFC data

DDD_LFC<- all_dolcetto_LFC_Gene %>% select(Gene,func, WT1D21=D2M1_All_D21, WT2D18=D2M2_All_D18, Mutant2D18=ARID1A_HT_All_D18)

p1<- DDD_LFC %>% ggplot(aes(y=WT1D21, x=Mutant2D18))+ geom_point()+stat_cor()
p2<- DDD_LFC %>% ggplot(aes(y=WT2D18, x=Mutant2D18))+ geom_point()+stat_cor()
p3<- DDD_LFC %>% ggplot(aes(y=WT1D21, x=WT2D18))+ geom_point()+stat_cor()
ggarrange(p1,p2,p3, nrow=1)
```

two wild type screens are correlated so we decided to use their mean for further analyses

```{r}
DDD_LFC$WT_av<- (DDD_LFC$WT1D21 +DDD_LFC$WT2D18)/2

### lets see the density plot
melted_ddd_lfc<- DDD_LFC %>% select(Gene, func, WT_av, Mutant2D18) %>% melt()

plot3<- melted_ddd_lfc %>% filter(func %in% c("CEG","NEG")) %>%  ggplot(aes(x=value, color= func, linetype=variable))+geom_density()+ theme_su()+scale_color_nejm(name="Gene Function", 
                                                                          labels=c("CEG"="Core Essential Genes" , "NEG"="Non-Essential Genes",  "Non-targeting_control"="Control", "Other"="Other Genes"))+
 xlab("Gene Mean LFC")+ ggtitle("Wild Type vs Mutant lines\nLFC")+ scale_linetype(name="Screen")


plot4<- DDD_melted_BF %>% filter(func %in% c("CEG","NEG"))%>% ggplot(aes(x=value, color= func, linetype=variable))+geom_density()+ theme_su()+scale_color_nejm(name="Gene Function", 
                                                                          labels=c("CEG"="Core Essential Genes" , "NEG"="Non-Essential Genes",  "Non-targeting_control"="Control", "Other"="Other Genes"))+
 xlab("BF")+ ggtitle("Wild Type vs Mutant lines\nBF")+ scale_linetype(name="Screen")

supp_fig_5_b<- ggarrange(plot3,plot4, ncol = 1, common.legend = T, legend = "right")
supp_fig_5_b

ggsave(supp_fig_5_b, filename= "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_5/supp_fig_5_b_new.svg", height = 4, width =4, units = "in", dpi=1000)

```

```{r}

DDD_LFC<- DDD_LFC %>% mutate(difference=(WT_av - Mutant2D18)) %>% mutate(scaled=scale(difference))

DDD_LFC<- DDD_LFC %>%  mutate(group="Normal") %>% mutate(group=case_when(
  difference > mean(difference) + 2*sd(difference) ~ "Depleted",
  difference < mean(difference) - 2*sd(difference) ~ "Enriched",
  TRUE ~as.character(group)))
table(DDD_LFC$group)


```


annotate enriched and depleted genes according to LFC


```{r}
DDD_LFC %>% ggplot(aes(x=WT_av,y=Mutant2D18, color=group))+
  geom_point()

```
differential essential genes are annotated differently with different analysis methods. as follow-up difference analysis will be replicated with a small pooled library

###small pooled library design

```{r}
bagel_selected_genes<- DDD_bagel %>% filter(group != "Normal")
lfc_seelcted_genes<- DDD_LFC %>% filter(group != "Normal")

DDD_bagel<- DDD_bagel %>% mutate(pool_group="none") %>%mutate(pool_group=case_when(
  Gene %in% bagel_selected_genes$Gene & Gene %in% lfc_seelcted_genes$Gene~ "both",
  Gene %in% bagel_selected_genes$Gene ~ "bagel",
  Gene %in% lfc_seelcted_genes$Gene~ "LFC",
  TRUE ~ as.character(pool_group)))


pool_genes<- DDD_bagel %>% filter(pool_group !="none") %>% droplevels()






```
## genes depleted or enriched by either of the methods are selected for follow-up. genes with non-specific guides and bidirectional promotes are eliminated:
a final pooled library is designed by adding 70 Non essential genes and 200 non-targeting control guides.



### end of analysis



