---
title: "GSEA Analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

1. gsea analysis: gene list pre-ranked by baysian factors (bf) + msigdb.v7.4 full gene sets

code:
gmxfile=msigdb.v7.4.symbols.gmt
rnkfile=ipsc_analyses_bf_values_wt_v_mut.rnk
~/programs/gsea/GSEA_4.1.0/gsea-cli.sh GSEAPreranked -gmx $gmxfile -collapse No_Collapse -mode Max_probe -norm meandiv -nperm 1000 -rnk $rnkfile -create_svgs false -include_only_symbols true -make_sets true -plot_top_x 20 -rnd_seed timestamp -set_max 500 -set_min 10 -zip_report false -out ipsc_bf_wt_v_mut_all



```{r}

library(tidyverse)
library(ggsci)
library(reshape2)
library(ggpubr)
library(cowplot)
library(svglite)

```

```{r}
theme_su<-  function(){ 
  font <- "Arial"   #assign font family up front
  
  theme_bw() %+replace%    #replace elements we want to change
    
    theme(
   
      
      #text elements
      plot.title = element_text(             #title
        family = "Arial",            #set font family
        size = 8,                #set font size
        face = 'bold',            #bold typeface
        hjust = 0,                #left align
        vjust = 3),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
        family = "Arial",            #font family
        size = 6),               #font size
      
      plot.caption = element_text(           #caption
        family = "Arial",            #font family
        size = 6,                 #font size
        hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
        family = "Arial",
        
        size = 6),               #font size
      
      axis.text = element_text(              #axis text
        family = "Arial",            #axis famuly
        size = 6),                #font size
      
      axis.text.x = element_text(            #margin for axis text
        margin=margin(5, b = 10)),
      
     legend.title=element_text(
       family="Arial",
       size=8,
       hjust = 0),
     legend.text=element_text(
       family = "Arial",
       size=6),
     
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank()
     )
  
     
    
}
```


```{r}

gsea<- read_tsv(file="input/Gene_set.txt")

```

```{r}

gsea<- gsea %>% mutate(data_base=case_when(
    grepl('GOBP', NAME)  ~ "GOBP",
    grepl('GOMF', NAME)  ~ "GOMF",
    grepl('GOCC', NAME)  ~ "GOCC",
    grepl('REACTOME', NAME)  ~ "REACTOME",
    grepl('KEGG', NAME)  ~ "KEGG"))

a<- gsea %>% filter(!is.na(data_base))
a<- a %>% filter(`FDR q-val`<0.05)
a<- arrange(a, 'ES')
b<- a %>% filter(data_base=="GOCC") %>% slice(1:20)
names(b)
b$NAME <- gsub(x = b$NAME, pattern = "GOCC_", replacement = "")
library(forcats)

fig_6_e<- b %>%  mutate(NAME = fct_reorder(NAME, ES))%>% ggplot(aes(x=`ES`, y=NAME,))+geom_point(color="#B22222")+theme_su()+ggtitle("Gene Set Enrichment Analysis")+ylab("Gene Ontology Cellular Component")+ xlab("Enrichment Score")
fig_6_e

c<- a %>% filter(data_base=="KEGG") %>% slice(1:20)

sup_fig_6_x<- c %>% ggplot(aes(x=`ES`, y=NAME, size=`FDR q-val`))+geom_point(color="#B22222")+theme_su()+ggtitle("Gene Set Enrichment Analysis")+ylab("KEGG Pathway")+ xlab("Enrichment Score")

sup_fig_6_x

d<- a %>% filter(data_base=="GOBP") %>% slice(1:20)

sup_fig_6_y<- d %>% ggplot(aes(x=`ES`, y=NAME, size=`FDR q-val`))+geom_point(color="#B22222")+theme_su()+ggtitle("Gene Set Enrichment Analysis")+ylab("Gene Ontology Biologic Process")+ xlab("Enrichment Score")

sup_fig_6_y

ggsave(fig_6_e, filename = "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_6/fig_6_e_new.svg", width = 5, height = 4, units = "in", dpi=1000)

ggsave(sup_fig_6_x, filename = "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_6/sup_fig_6_e.svg", width = 4, height = 4, units = "in", dpi=1000)
ggsave(sup_fig_6_y, filename = "E:/Paper writing/CRIPSRi paper/Analyses/plots/panel_6/sup_fig_6_f.svg", width = 7, height = 4, units = "in", dpi=1000)

```


