---
title: "activity_plotting"
author: "Sunay"
date: '2022-05-20'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggsci)
library(tidyverse)

```
```{r}
theme_su<-  function(){ 
  font <- "Arial"   #assign font family up front
  
  theme_bw() %+replace%    #replace elements we want to change
    
    theme(
   
      
      #text elements
      plot.title = element_text(             #title
        family = "Arial",            #set font family
        size = 8,                #set font size
        face = 'bold',            #bold typeface
        hjust = 0,                #left align
        vjust = 3),               #raise slightly
      
      plot.subtitle = element_text(          #subtitle
        family = "Arial",            #font family
        size = 6),               #font size
      
      plot.caption = element_text(           #caption
        family = "Arial",            #font family
        size = 6,                 #font size
        hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
        family = "Arial",
        
        size = 6),               #font size
      
      axis.text = element_text(              #axis text
        family = "Arial",            #axis famuly
        size = 6),                #font size
      
      axis.text.x = element_text(            #margin for axis text
        margin=margin(5, b = 10)),
      
     legend.title=element_text(
       family="Arial",
       size=8,
       hjust = 0),
     legend.text=element_text(
       family = "Arial",
       size=6),
     
     panel.grid.major = element_blank(),
     panel.grid.minor = element_blank()
     )
  
     
    
}
```

## read in results of all activity tests
Activity test is conducted using a reporter plasmid expressing GFP and BFP from different promoters and as well as a gRNA targeting GFP promoter (or a mock gRNA). The activity is determined as ercent decrease in median GFP signal in BFP positive cells transfected with the targeting gRNA compared to cells transfected with non-targeting (mock) gRNA. 
```{r}

all_activity_data<- read_csv(file="input/all_activity.txt")
names(all_activity_data)

summary<- all_activity_data %>% group_by(Clonality, Construct, Line) %>% summarise(Mean=mean(Activity),Median=median(Activity), Sd=sd(Activity))
write_csv(summary, file="input/activity_summary.txt")

summary %>% ggplot(aes(x=Line, fill=Construct, y=Mean))+ geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=Mean-Sd, ymax=Mean+Sd), width=.2,
                 position=position_dodge(.9)) +facet_wrap(~Clonality)

all_activity_data<- all_activity_data %>% unite("Cons", Clonality:Construct, remove = F)

ipsc_activity<- all_activity_data %>% filter(Line=="hIPSC") %>% droplevels()


figure_1_b<- ipsc_activity %>% ggplot(aes(x=Cons, y=Activity, fill=Construct, linetype=Clonality))+geom_boxplot()+theme_su()+theme(axis.text.x = element_blank(), legend.position = "none")+ylab("Percent dCas9 Activity")+ggtitle("dCas9 Activity in hIPSC Cells")+xlab("Cell Lines")+scale_fill_nejm(name="Cell Lines")
figure_1_b

svg(filename= "./plots/figure_1_b.svg", width = 3, height = 2)
plot(figure_1_b)
dev.off()

figure2<- all_activity_data %>% ggplot(aes(x= Cons, y=Activity, fill=Line))+geom_boxplot()+theme_su()+theme(axis.text.x = element_text(angle=90), axis.title = element_text(face="bold"))+scale_fill_nejm(name="Cell Line")+xlab("dCas9 fusion constructs")+ggtitle("Activity of dCas9 Fusion Proteins in hiPSC and K562 Lines")
figure2

svg(filename= "./plots/supplementary_figure_1.svg", height = 5, width = 10)
plot(figure2)
dev.off()
ggsave(figure2, filename= "./plots/supplementary_figure_1.jpeg", dpi=1000, width = 10, height = 5, units = "cm")

```


